### 2017-03-15 paper/main.tex
Started on paper

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr} \usepackage{times} \usepackage{epsfig}
\usepackage{graphicx} \usepackage{amsmath} \usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[breaklinks=true,bookmarks=false]{hyperref}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{1} \begin{document}

%%%%%%%%% TITLE
\title{\LaTeX\ Author Guidelines for CVPR Proceedings}

\author{Weilian Song\\ 
University of Kentucky\\ 
{\tt\small
weilian.song@uky.edu}
% For a paper whose authors are all at the same institution, omit the
% following lines up until the closing ``}''.  Additional authors and
% addresses can be added with ``\and'', just like the second author.  To
% save space, use either the email address or home page, not both
\and Nathan Jacobs\\ 
University of Kentucky\\ 
{\tt\small jacobs@cs.uky.edu}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract} TODO: Make Abstract last \end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}

Acoustic detection of birds in the wild is useful in many biological
scenarios, including estimation of bird population and species-specific
acoustic monitoring. Most of the current methods are either manual or
semiautomatic, requireing labor-intensive processing/tuning.

In response to the need for automatic, tuning-free bird presentce detection
algorithm, Bird Audio Detection Challenge was created, hosted by Queen Mary
University of London and other universities. Three datasets were released
as part of the challenge, and teams across the world compete for the best
performance on the held-out dataset.

In section 2, we will discuss some of the competition details, including
the datasets used, task description and the evaluation metrics. Section 3
will present three methods for bird acoustic detection, two of which are
our submission to the challenge, and the third one being the best
performing method submitted by Thomas Grill. Section 4 will discuss the
evaluation results of the three methods, and Section 5 will present future
possible work based off of this challenge.

%------------------------------------------------------------------------
\section{Competition Details}

The Bird Audio Detection Challenge is hosted by Queen Mary University of
London, in collaboration with the IEEE Signal Processing Society.

%-------------------------------------------------------------------------
\subsection{Task Description}

For the challenge, teams have to design a system that takes a short audio
clip, and returns a binary label denoting the presence/absence of bird
sound of any kind. Teams are also encouraged to output a continuous,
probability score in the range of 0 to 1 for evaluation purposes.

%-------------------------------------------------------------------------
\subsection{Datasets}

Three datasets are provided as part of the challenge.

\textbf{freefield1010:} 7000 excerpts from field recordings around the
world, gathered by the FreeSound project.

\textbf{Warblr:} Crowdsourced project with many smartphone recordings from
around the United Kingdom. Very diverse in sound categories, even including
human bird imitations.

\textbf{Chernobyl:} Dataset created as part of Dr Wood's research into the
long-term effects of the Chernobyl accident on local ecology. Approximately
10,000 hours to date

Each dataset contains audio clips of roughly 10 seconds. For the challenge,
all datasets' audio files are provided, but only \textbf{freefield} and
\textbf{Warblr}'s labels are given. Held-out dataset consists primarily of
the \textbf{Chernobyl} dataset, along with some from the other two.

%-------------------------------------------------------------------------
\subsection{Evaluation Metrics}

To evaluate the performance of a model, we use the standard "Area Under the
ROC Curve" method, also known as the AUC score.

%-------------------------------------------------------------------------
\section{Bird Acoustic Detection}

Here we introduce three methods for bird acoustic detection, two of which
are proposed by us, and the third one being the best performing method by
Thomas Grill of University of Music and Performing Arts, Vienna.

%-------------------------------------------------------------------------
\subsection{Data preparation and augmentation}

All audio clips are 10 seconds, mono at 44100 Hz, which gives us 441000
frames of audio. We divide all frames by 32768 to convert values of
individual frames to between -1 and 1, and randomly crop out 400000 frames
as input to our network.

We also deployed data augmentation techniques. Given one base sound clip,
positive or negative, and another negative sound clip, we concatenate them
along the y-axis, with the base clip weighted by a factor \(r\) between 0.25
and 1, and the negative sound clip weighted by a factor \(1-r\). So we
have:
\begin{equation}
	x_{augmented}=r*x_{signal} + (1-r)*x_{negative}
\end{equation}
If \(r\) equals to 1, then there would be no augmentation, and if \(r\)
equals to 0.25, only 25\% of the base sound clip will be preserved. We
prevent \(r\) from going below 0.25 to ensure bird sound is still
detectable after augmentation.

Label for the augmented clip will remain the same as the base clip, as the
augmentation is an \(|\) process, preserving the original label when \(|\)
with a negative label.

%-------------------------------------------------------------------------
\subsection{Method 1}

Our first method resembles basic characteristics of SoundNet and residual
net. The network is fully convolutional, with one-dimensional convolutions
followed by nonlinearities (ReLU) layers. This network is suitable for
audio signals, as it is invariant to translation and allows stacking of
convolutional layers for detecting high-level features.

Network also has shortcuts from earlier layers of the network to later
stages, allowing easier gradient back-propagation to early layers of the
network. See Table 1 for network details.

%------------------------------------------------------------------------

\subsection{Method 2}

Our second method utilizes a series of pooling and convolutional layers to
extract sets of features from the input audio at different resolutions, and
concatenate them together before some final convolutional layers.

Network can be divided into three stages. In the first stage, we first
create two copies of the input, one reshaped to 100000 by 4, and the other
reshaped to 200000 by 2. \(x_{100000*4}\) is then passed to a convolutional
layer with kernel size \([3,4]\) while \(x_{200000*2}\) is passed to a
series of 5 pooling layers, followed by a convolutional layer at the end of
each pooling layer. Note here that the input of one pooling layer is the
output of the previous pooling layer, the outputs of the convolutional
layers serve as features extracted at audio of different resolution.

In the second stage, we combined the outputs of the 6 convolutional layers
in Stage 1 into one feature vector. First, we find the maximum for each
convolutional output along the thrid dimension. We denote these as the
\textit{energies} for each convolutional layer. We then concatenate them
with their corresponding convolutional output, run each concatenated
feature vector through a pooling layer of specified size, and concatenate
all 6 feature vectors into one along the third dimension. Due to different
lengths of output for the 6 convolutional layers, we pass them through
pooling layers of different sizes to ensure that they are of the same
dimension for concatenation.

In the final stage, we pass the output from Stage 2 through 7 additional
convolutional layers, with a \(max()\) operation between the 4th and 5th
layer. Final output is then flattened into size (1,2), ready to be used in
a softmax classification loss function.

%------------------------------------------------------------------------

\subsection{Method 3}

This method is the winning algorithm, submitted by Thomas Grill.

%------------------------------------------------------------------------
\section{Final copy}

You must include your signed IEEE copyright release form when you submit
your finished paper. We MUST have this form before your paper can be
published in the proceedings.

{\small \bibliographystyle{ieee} \bibliography{biblio} }

\end{document}