### 2017-03-17 paper/main.tex
Added figures and finished most of writing

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr} \usepackage{times} \usepackage{epsfig}
\usepackage{graphicx} \usepackage{amsmath} \usepackage{amssymb}

\graphicspath{ {figures/} }

\newenvironment{alt_enumerate}
{ \begin{enumerate}
		\setlength{\itemsep}{1pt}
		\setlength{\parskip}{1pt}
		\setlength{\parsep}{1pt} }
	{ \end{itemize} }

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[breaklinks=true,bookmarks=false]{hyperref}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{1} \begin{document}

%%%%%%%%% TITLE
\title{Bird Audio Detection Challenge}

\author{Weilian Song\\ 
University of Kentucky\\ 
{\tt\small
weilian.song@uky.edu}
% For a paper whose authors are all at the same institution, omit the
% following lines up until the closing ``}''.  Additional authors and
% addresses can be added with ``\and'', just like the second author.  To
% save space, use either the email address or home page, not both
\and Nathan Jacobs\\ 
University of Kentucky\\ 
{\tt\small jacobs@cs.uky.edu}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}

	Acoustic detection of birds in the wild is useful in many biological
	scenarios, including estimation of bird population and species-specific
	acoustic monitoring. Most of the current methods are either manual or
	semiautomatic, requiring labor-intensive processing/tuning. We hereby
	present three alternative methods, as part of the Bird Audio Detection
	Challenge, hosted by Queen Mary University of London. Two methods are of
	our own submission, with the third one as the winning solution submitted
	by Thomas Grill. All methods utilize deep learning for automatic,
	tuning-free bird presence detection.

\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}

Acoustic detection of birds in the wild is useful in many biological
scenarios, including estimation of bird population and species-specific
acoustic monitoring. Most of the current methods are either manual or
semiautomatic, requireing labor-intensive processing/tuning.

In response to the need for automatic, tuning-free bird presentce detection
algorithm, Bird Audio Detection Challenge was created, hosted by Queen Mary
University of London and other universities. Three datasets were released
as part of the challenge, and teams across the world compete for the best
performance on the held-out dataset.

In section 2, we will discuss some of the competition details, including
the datasets used, task description and the evaluation metrics. Section 3
will present three methods for bird acoustic detection, two of which are
our submission to the challenge, and the third one being the best
performing method submitted by Thomas Grill. Section 4 will discuss the
evaluation results of the three methods, and Section 5 will present future
possible work based off of this challenge.

%------------------------------------------------------------------------
\section{Competition Details}

The Bird Audio Detection Challenge is hosted by Queen Mary University of
London, in collaboration with the IEEE Signal Processing Society.

%-------------------------------------------------------------------------
\subsection{Task Description}

For the challenge, teams have to design a system that takes a short audio
clip, and returns a binary label denoting the presence/absence of bird
sound of any kind. Teams are also encouraged to output a continuous,
probability score in the range of 0 to 1 for evaluation purposes.

%-------------------------------------------------------------------------
\subsection{Datasets}

Three datasets are provided as part of the challenge.

\textbf{freefield1010:} 7000 excerpts from field recordings around the
world, gathered by the FreeSound project.

\textbf{Warblr:} Crowdsourced project with many smartphone recordings from
around the United Kingdom. Very diverse in sound categories, even including
human bird imitations.

\textbf{Chernobyl:} Dataset created as part of Dr Wood's research into the
long-term effects of the Chernobyl accident on local ecology. Approximately
10,000 hours to date

Each dataset contains audio clips of roughly 10 seconds. For the challenge,
all datasets' audio files are provided, but only \textbf{freefield} and
\textbf{Warblr}'s labels are given. Held-out dataset consists primarily of
the \textbf{Chernobyl} dataset, along with some from the other two.

%-------------------------------------------------------------------------
\subsection{Evaluation Metrics}

To evaluate the performance of a model, we use the standard "Area Under the
ROC Curve" method, also known as the AUC score.

%-------------------------------------------------------------------------
\section{Bird Acoustic Detection}

Here we introduce three methods for bird acoustic detection, two of which
are proposed by us, and the third one being the best performing method by
Thomas Grill of University of Music and Performing Arts, Vienna.

%-------------------------------------------------------------------------
\subsection{Data preparation}

All audio clips are 10 seconds, mono at 44100 Hz, which gives us 441000
frames of audio. We divide all frames by 32768 to convert values of
individual frames to between -1 and 1, and randomly crop out 400000 frames
as input to our network.

%-------------------------------------------------------------------------
\subsection{Method 1}

Our first method resembles basic characteristics of SoundNet and residual
net. The network is fully convolutional, with one-dimensional convolutions
followed by nonlinearities (ReLU) layers. This network is suitable for
audio signals, as it is invariant to translation and allows stacking of
convolutional layers for detecting high-level features.

Network also has shortcuts from earlier layers of the network to later
stages, allowing easier gradient back-propagation to early layers of the
network. See Figure 1 for network details.

\begin{figure}[t]
	\centering
	\includegraphics[width=0.5\textwidth]{v5_diagram}
	\caption{Network v5 Architecture. Content with box refers to kernel size,
	number of filters, and stride size}
\end{figure}

%------------------------------------------------------------------------

\subsection{Method 2}

Our second method utilizes a series of pooling and convolutional layers to
extract sets of features from the input audio at different resolutions, and
concatenate them together before some final convolutional layers.

Network can be divided into three stages. In the first stage, we first
create two copies of the input, one reshaped to 100000 by 4, and the other
reshaped to 200000 by 2. \(x_{100000*4}\) is then passed to a convolutional
layer with kernel size \([3,4]\) while \(x_{200000*2}\) is passed to a
series of 5 pooling layers, followed by a convolutional layer at the end of
each pooling layer. Note here that the input of one pooling layer is the
output of the previous pooling layer, the outputs of the convolutional
layers serve as features extracted at audio of different resolution.

In the second stage, we combined the outputs of the 6 convolutional layers
in Stage 1 into one feature vector. First, we find the maximum for each
convolutional output along the thrid dimension. We denote these as the
\textit{energies} for each convolutional layer. We then concatenate them
with their corresponding convolutional output, run each concatenated
feature vector through a pooling layer of specified size, and concatenate
all 6 feature vectors into one along the third dimension. Due to different
lengths of output for the 6 convolutional layers, we pass them through
pooling layers of different sizes to ensure that they are of the same
dimension for concatenation.

In the final stage, we pass the output from Stage 2 through 7 additional
convolutional layers, with a \(max()\) operation between the 4th and 5th
layer. Final output is then flattened into size (1,2), ready to be used in
a softmax classification loss function.

\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{v8_diagram}
	\caption{Network v8 Architecture. Content of convolutional blocks refer
	to kernel size, number of filters, and stride size. Content of avgpool
	blocks refer to kernel size and stride size.}
\end{figure*}

%------------------------------------------------------------------------
\subsection{Method 3}

% TODO: Explain that spectrograms were used during training

This method is the winning algorithm of the challenge, submitted by Thomas
Grill. The neural network used consists of 4 convolutional layers with
pooling, and 2 dense layers. Multiple signal pre-processing techniques were
used, including silence/noise trimming, and a different data augmentation
technique involving shifting the spectra in time and frequency.
Pseudo-labeling and bagging techniques are also deployed, which we will
discuss in section 4.

%------------------------------------------------------------------------
\section{Experiments}

During our development and training of the network, we experimented with
various pre- and postprocessing techniques, as well as some training
strategies.

%------------------------------------------------------------------------
\subsection{Implementation Details}

The given labeled data from both dataset is split into 10 folds. Within
each fold, number of positive and negative samples are balanced. During
training, number of samples loaded from each dataset are also balanced, to
ensure no bias.

For in-house testing, 9 folds are used for training and 1 folf for testing.
For submission, all 10 folds are used for training.

For our method 1 and 2, we implemented them in python with Tensorflow
machine learning library. We trained from scratch, with learning rate of
0.03, learning rate decay by a factor of 0.1 every 40000 iterations, and a
total of 300000 iterations.

%------------------------------------------------------------------------
\subsection{Data Augmentation}

We deployed data augmentation as our only pre-processing technique. Given
one base sound clip, positive or negative, and another negative sound clip,
we concatenate them along the y-axis, with the base clip weighted by a
factor \(r\) between 0.25 and 1, and the negative sound clip weighted by a
factor \(1-r\). So we have:
\begin{equation}
	x_{augmented}=r*x_{signal} + (1-r)*x_{negative}
\end{equation}
If \(r\) equals to 1, then there would be no augmentation, and if \(r\)
equals to 0.25, only 25\% of the base sound clip will be preserved. We
prevent \(r\) from going below 0.25 to ensure bird sound is still
detectable after augmentation.

Label for the augmented clip will remain the same as the base clip, as the
augmentation is an \(|\) process, preserving the original label when \(|\)
with a negative label.

%------------------------------------------------------------------------
\subsection{Reducing capacity}

One main difficulty we faced is the small size of training data. With no
regularization, our networks consistently over-fit to the training data,
perform near perfect on training and our own hold-out test data, but
perform poorly on the challenge data. To compensate for the lack of data,
we reduce the capacity of our models by reducing the number of filters for
each convolutional layer. We do so until the network cannot perform
perfectly on the training and testing data.

%------------------------------------------------------------------------
\subsection{Prediction Bagging}

This post-processing technique averages prediction scores from many models.
Due to the random cropping of audio signals during the data loading phase,
different evaluation of the same audio clip will yield slightly different
results. Therefore we predict 100 sets of scores for the held-out challenge
dataset, and average them to obtain one set of scores that was submitted.

%------------------------------------------------------------------------
\subsection{Domain Confusion}

% TODO Consult with Dr. Jacobs on whether this should be mentioned or not

This training strategy addresses the issue of dataset bias. It is apparent
that the three given datasets have disjoint domains. In addition, within
each dataset, data is skewed either towards mostly positive samples or
mostly negative ones. During training, it is very likely that the network
will cheat by identifying which dataset the clip is from, and make its
prediction based off of that information. In response, we take the audio
features near the end of the network, pass it through another convolutional
layer, and...

%------------------------------------------------------------------------
\subsection{Pseudo Labeling}

This technique is used by many contestants and attempted by us. The
assumption is that high-confidence predictions on the held-out challenge
dataset are accurate enough to be used as extra data for retraining or
fine-tuning. Basic procedures are:

\begin{alt_enumerate}
	\item{Train a model with only labeled data}
	\item{Predict scores of the held-out challenge dataset}
	\item{Find threshold where within it, predictions are accurate enough to
		be used as extra labels}
	\item{Add all predictions within threshold, and retrain or fine-tune
		network}
\end{alt_enumerate}

%------------------------------------------------------------------------
\section{Results}

Below are the final AUC scores for each of the 3 methods:\\

\begin{tabular}{ |p{2cm}|p{2cm}|p{2cm}|p{2cm}| }
	\hline
	\multicolumn{2}{|c|}{AUC Scores}\\
	\hline
	Method & AUC Score\\
	\hline
	Method 1 & 77.9\\
	Method 2 & ??.?\\
	Method 3 & 88.7\\
	\hline
\end{tabular}

%------------------------------------------------------------------------
\section{Future Work}

For method 1 and 2, no signal pre-processing techniques were deployed,
resulting in very noisy inputs to our network. It will be interesting to
find out if pre-processing will aid the prediction accuracy of our network.

%------------------------------------------------------------------------
\section{Conclusion}

Bird acoustic detection is an important task for many biological
applications, and the challenge has advanced this field in many ways. We
presented three methods, first two with a focus on unique network
architecture, and the third one as the embodiment of many years of machine
listening research by Thomas Grill. Many topics are still open for
research, like bird species identification and detection.

%------------------------------------------------------------------------

{\small \bibliographystyle{ieee} \bibliography{biblio} }

\end{document}